\section{Reinforcement Learning Approaches for Personalized Applications}

Reinforcement Learning (RL) is a powerful method for building personalized systems. It helps systems to learn and improve based on user interactions over time. Many RL techniques have been used in different fields like entertainment, healthcare, education, and e-commerce. This section explains some popular RL approaches that are used for personalization.

\paragraph*{Contextual Bandits} This method is used by Netflix to show different artwork (thumbnails) for movies and series to different users. It learns from the user's past behaviour and decides which artwork will attract the user more. It balances between trying new options (exploration) and using known successful options (exploitation) \cite{netflix2018artwork}.

\paragraph*{Proximal Policy Optimization (PPO)} In healthcare, PPO has been used to suggest personalized cancer treatments. It works by ranking different drugs based on the data from each patient. This helps doctors to choose better treatments that match each patient's unique condition \cite{liu2022deep}.

\paragraph*{Deep Q-Learning (DQN)} DQN has been applied to recommend projects to users based on their interests. It improves the accuracy of recommendations and helps users to trust the system more \cite{wang2019personalized}.

\paragraph*{Q-Learning} This is a simple and popular RL method. It can be used for many general recommendation tasks and is good at adapting to changes in user preferences \cite{edirisinghe2020reinforcement}.

A study done by Moise et al. \cite{moise2020reinforcement} reviewed 166 research papers and found that Q-learning is the most commonly used RL method (used in 60 studies). It has been used in many areas like healthcare, entertainment, education, and commerce. According to the study, commerce and entertainment domains often use realistic experiments, while healthcare and communication have limitations due to safety and data issues.

\begin{table}[htb]
\centering
\small
\begin{tabular}{|p{2.8cm}|p{3cm}|p{2.8cm}|p{4cm}|}
\hline
\textbf{Approach} & \textbf{Application} & \textbf{Domain} & \textbf{Key Features} \\
\hline
Contextual Bandits & Artwork Personalization & Entertainment & Balances exploration/exploitation, scalable \\
\hline
Proximal Policy Optimization (PPO) & Treatment Recommendation & Healthcare & Ranks drugs using DRL, handles high-dimensional data \\
\hline
Deep Q-Learning (DQN) & Project Recommendations & Recommendation & Learns user preferences, boosts trust \\
\hline
Q-Learning & General Recommendations & Recommendation & Adapts to changing preferences, versatile \\
\hline
\end{tabular}
\caption{Summary of Reinforcement Learning Approaches for Personalized Applications}
\label{tab:rl-approaches}
\end{table}
