\documentclass[10pt,a4paper]{article}

% PACKAGES
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm,paperwidth=170mm,paperheight=240mm]{geometry} % Approximates 122mm x 193mm text area
\usepackage{fancyhdr}
\usepackage{parskip}
\usepackage{natbib}
\usepackage{imakeidx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage[acronym]{glossaries}
\usepackage{lipsum} % DUMMY TEXT
\usepackage[english]{babel} % Assuming English
\usepackage{titlesec}
\usepackage{tabularx}
\usepackage{float}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{array}
\usepackage{listings}
\usepackage{placeins}
\usepackage{longtable}
\usepackage{color}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{makecell}

% CONFIGS
\pagestyle{plain}
\fancyfoot[C]{\thepage}
\hypersetup{
    pdfauthor = {Avishka Sathyanjana},
    pdftitle = {Multimodal Emotional State Recognition for Personalized Responses},
    colorlinks,
    linkcolor={black},
    citecolor={black}
}
\PassOptionsToPackage{unicode}{hyperref}

% MACROS
\newcommand{\unit}[1]{\ensuremath{\, \mathrm{#1}}}  % Makes writing units simpler
\renewcommand{\thefootnote}{\emph{\alph{footnote}}} % Italic letters for footnotes

% PATHS
\graphicspath{ {img/} }

% BIB SETUP
\bibliographystyle{plainnat} % Numerical citation style

% TITLE FORMAT
\title{\textbf{\fontsize{14}{16}\selectfont Multimodal Emotional State Recognition for Personalized Responses}}
\author{
    \fontsize{10}{12}\selectfont Avishka Sathyanjana \\
    \fontsize{9}{11}\selectfont University of Moratuwa, Sri Lanka \\
    \fontsize{9}{11}\selectfont avishka.sathyanjana@example.com
}

% HEADINGS FORMAT
\titleformat{\section}{\normalfont\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\normalfont\normalsize\bfseries}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}

% DOCUMENT BEGIN HERE
\begin{document}

% TITLE AND AUTHOR
\maketitle

% ABSTRACT
\begin{abstract}
    \fontsize{9}{11}\selectfont
  This study presents a personalized multimodal emotion recognition framework that adapts to individual emotional baselines using facial and vocal signals. Decision Level Fusion with Mean Squared Error assigns personalized weights based on user-reported emotions. Kernel Density Estimation establishes an initial arousal-valence baseline, refined via reinforcement learning with emoji-based feedback. Experiments with 10 participants across five emotions (Happy, Angry, Sad, Boredom, Calm) show the fused method improves accuracy by 33.92\% over facial and 6.52\% over vocal methods. Personalized responses enhanced empathy (75.3\%), emotional alignment (69.5\%), and satisfaction (37.6\%). Most participants (66.67\%) validated the refined baseline. Contributions include a personalized fusion method, baseline identification, and iterative refinement. Future work will expand emotional categories, incorporate physiological data, and enhance context-aware fusion. This framework advances personalized human-computer interaction.
\end{abstract}

% KEYWORDS
\textbf{Keywords:} Affective Computing, Emotional State Recognition, Multimodal Analysis, Personalized Responses, Machine Learning, Human-Computer Interaction, Reinforcement Learning, Arousal-Valence, 

% INTRODUCTION
\input{src/chapter_01/index.tex}

% LITERATURE REVIEW
\section{Literature Review}
\input{src/chapter_02/index.tex}

% DESIGN AND IMPLEMENTATION
\section{Design and Implementation}
\input{src/chapter_03/index.tex}

% RESULTS AND EVALUATION
\section{Results and Evaluation}
\input{src/chapter_04/index.tex}

% CONCLUSION
\section{Conclusion}
\input{src/chapter_05/index.tex}

% ACKNOWLEDGMENTS
\section*{Acknowledgments}
The authors would like to thank the University of Moratuwa for their support.

% CRediT AUTHOR STATEMENT
\section*{CRediT author statement}
Avishka Sathyanjana: Conceptualization, Methodology, Formal analysis, Writing – original draft preparation.

% REFERENCES
\newpage
\bibliography{refernce/references}

\end{document}